Discrete Quantum Theories and Computing

### 2018-12-01 dissertation.tex
From: Hanson, Andrew J. <hansona@indiana.edu> Sent: Thursday, November 29, 2018 13:24 To: Tai, Yu-Tsung <yuttai@indiana.edu> Cc: Ortiz, Gerardo <ortizg@indiana.edu>; Sabry, Amr A. <sabry@indiana.edu> Subject: CV...abstract edits...

\begin{abstract*}\todo{Rewrite the abstract so that the committee
can sign it. Prof. Thurston: ``To address the same issue, the thesis
contains four different theories each have different pros and cons.
The abstract should said them seperately and explicitly.'' }Most
of the quantum computing models are based on the continuum of real
numbers, despite the fact that classical digital computers faithfully
realize the prediction of discrete computational models, while analog
computers in reality are far weaker than the theoretical prediction
of continuous computational models. Since one of the important factors
of this gap is the measurement precision, we hope to build a more
faithful model by investigating discrete quantum computing models
and incorporating the idea of finite precision measurement into the
quantum theory.

We first replace the continuum of complex numbers by discrete finite
fields in the quantum theory. The simplest theory, defined over unrestricted
finite fields, is so weak that it cannot express Deutsch's algorithm.
This quantum theory is, however, also so powerful that it can be used
to solve the $\usat$ problem, which is as hard as a general $\NP$-complete
problem. The second framework employs only finite fields with no solution
to $x^{2}+1=0$, and thus permits an elegant complex representation
of the extended field by adjoining $\rmi=\sqrt{-1}$. Because the
states of a discrete $n$-qubit system are in principle enumerable,
we can count the number of states, and determine the proportions of
entangled and unentangled states. Although this improved framework
allows the probabilistic Grover search algorithm in a local region,
we haven't found sensible correspondences between quantum probability
measures and the whole state space in general.

To search a sensible correspondence and to incorporate the idea of
finite precision measurement at the same time, we shift our attention
to consider quantum interval-valued probability measures (IVPMs).
This interval-valued framework not only provides a natural generalization
of both classical IVPMs and conventional quantum probability measures,
but also establishes bounds on the validity of the Kochen-Specker
and Gleason theorems in realistic experimental environments.\end{abstract*}

>>>>

\begin{abstract*}\todo{Rewrite the abstract so that the committee
can sign it. Prof. Thurston: ``To address the same issue, the thesis
contains four different theories each have different pros and cons.
The abstract should said them separately and explicitly.'' }

Most quantum computing models are based on the continuum of real numbers,
while classical digital computers faithfully realize only discrete
computational models. Analog computers appear to be an option, but
in reality are far weaker than would be needed for computational models
requiring real numbers. One approach to resolving this conflict is
to find consistent mathematical ways to limit measurement precision
to computable contexts that do not require incomputable real numbers.
Our goal is to build a more philosophically consistent models by investigating
discrete quantum computing using finite number systems, and, alternatively,
by incorporating finite precision measurement using intervals into
quantum theory.

We begin by replacing the continuum of complex numbers by discrete
finite fields in quantum theory. The simplest theory, defined over
unrestricted finite fields, is so weak that it cannot express Deutsch's
algorithm, but, paradoxically, is also so powerful that it can be
used to solve the $\usat$ problem, which is as hard as a general
$\NP$-complete problem. Our second framework employs only finite
fields over prime numbers of the form $4n+3$, which possess no solutions
to $x^{2}+1=0$, and thus permit an elegant complex representation
of the finite field by adjoining $\rmi=\sqrt{-1}$. Because the states
of a discrete $n$-qubit system are in principle enumerable, we can
count the number of states, and determine the proportions of entangled
and unentangled states. Although this improved framework can be used
to implement the probabilistic Grover search algorithm in a local
region, we haven't found a consistent way to treat quantum probability
measures in general.

Finally, we shift our attention to consider quantum interval-valued
probability measures (IVPMs), which potentially embody both finite
precision measurement and a sensible correspondence to standard quantum
probability. This interval-valued framework not only provides a natural
generalization of both classical IVPMs and conventional quantum probability
measures, but also establishes bounds on the validity of the Kochen-Specker
and Gleason theorems in realistic experimental environments.\end{abstract*}
